\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing economic inequality with tax data - Switzerland from 1945 to 2010}

### 2014-07-03 paper/subdocuments/results.Rnw

@@

<<setup_gini_data>>=
library(foreign,quietly=TRUE, warn.conflicts=FALSE)
library(ggplot2)
df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
df$zeros <- df$null_norm/df$cpop # i only use normal cases here
@

@@

<<setup_data_Brühlhart with and without taxed>>=
library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
# Ohne Nuller
bd <- read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd$Jahr <- as.numeric(substr(bd$Veranlagungsperiode,1,4))
bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] <- bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] - 2
bd<-bd %.% 
  filter(Einheit=="Total")
bd<-subset(bd,Jahr<1995 | Jahr >2002)
bd.long<-reshape(bd,
varying = c("gini_reink","gini_reinka"),
v.names = "Gini",
timevar = "inc_def",
times = c("Net income (only taxed)","Equivalised net income (only taxed)"),
direction ="long")
# mit Nuller
bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
bd.0<-bd.0 %.% 
  filter(Einheit=="Total")
bd.0<-subset(bd.0,Jahr<1993 | Jahr >2002)
bd.0.long<-reshape(bd.0,
varying = c("gini_reink","gini_reinka"),
v.names = "Gini",
timevar = "inc_def",
times = c("Net income","Equivalised net income"),
direction ="long")
bdlong<-rbind(bd.0.long,bd.long)
names(bdlong)[61]<-"Income"
@

<<equivalencescale, fig.cap='',fig.pos='H'>>=
library(ggplot2)
number_ticks <- function(n) {function(limits) pretty(limits, n)}
ggplot(bdlong, aes(x=Jahr,y=Gini,shape=Income))+
  scale_y_continuous(limits=c(0.3,0.55))+
  scale_x_continuous(limits=c(1970,2012),breaks=number_ticks(10)) +
  geom_point(aes(shape=Income),size=3)+
  geom_line(data=bdlong,aes(linetype=Income))+
  theme_bw()+
  xlab("tax period")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))
@

@@

<<setup_data_Brülhart_2003>>=
library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
library(foreign)
bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
bd.2003<-bd.0 %.% 
  filter(Jahr==2003,Einheit=="Total")
start <- which(names(bd.2003)=="p1")
end <- which(names(bd.2003)=="p99_99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
bd.2003<- data.frame(t(bd.2003[1,start:end]),percentile)
names(bd.2003)[1]=c("inc")
cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
bd.2003 <- rbind(c(0,0),bd.2003)
ival<-bd.2003$inc
perval<-bd.2003$percentile
bd.2003.ind<-unlist(sapply(2:length(ival),function(i){
  len=(perval[i]-perval[i-1])*100
  seq(ival[i-1],ival[i],length.out=len)
}))
bd.2003.ind<-bd.2003.ind/1000

@@

<<setup_data_Brülhart_2010>>=
library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
library(foreign)
bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
bd.2010<-bd.0 %.% 
  filter(Jahr==2010,Einheit=="Total")
start <- which(names(bd.2010)=="p1")
end <- which(names(bd.2010)=="p99_99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
names(bd.2010)[1]=c("inc")
bd.2010 <- rbind(c(0,0),bd.2010)
ival<-bd.2010$inc
perval<-bd.2010$percentile
bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
  len=(perval[i]-perval[i-1])*100
  seq(ival[i-1],ival[i],length.out=len)
}))
bd.2010.ind<-bd.2010.ind/1000
@

<<reldist20032010, echo=FALSE, fig.cap='',fig.pos='H'>>=
library(reldist, quietly=TRUE, warn.conflicts = FALSE)

@@

<<poltable>>=
# Setup functin to produce latex tables with polarization indices and difference in ginis
polTable <- function(y, yo, ywgt=1, yowgt=1,ndig=2,title="Inequality Indices") {
stopifnot(is.numeric(y), is.numeric(yo)) # all TRUE
#y vector with new distribution
#yo vector with reference distribution
#ywgt yowgt: weights to use (optional)
#ndig: digits to display, default 2
#title: table title (optional)

# load neccessary libraries
library(reldist,quietly=TRUE,warn.conflicts=FALSE) #for Latex tables
library(Hmisc,quietly=TRUE,warn.conflicts=FALSE)
library(xtable,quietly=TRUE,warn.conflicts=FALSE)

@@

latex(format(c("Median Index"=med, "Lower Index"=lower, "Upper Index"=upper, "$\\Delta$ Gini"=dgini),digits=ndig),file="",title=title,table.env=FALSE)
}

<<polindex20032010, results='asis'>>=
polTable(bd.2010.ind,bd.2003.ind)
@

@@

% Variante 1993/1994
<<setup_data_ESTVspecialcases9394>>=
library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly=TRUE, warn.conflicts = FALSE)
#library(reldist)
normal <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
normal <- select(normal, p1:p99, kanton, steuerperiode, -p96,-p97,-p98)
#normal <- normal %.% select(-G_reink, -G_taxed, -ppop0, -G_steink0)
pooled <- read.csv("../../data/ginis_und_perzentile_beides.csv",sep="\t")
pooled <- select(pooled, kanton, steuerperiode, p1:p99)
#special <- read.csv("data/ginis_und_perzentile_sonder.csv",segep="\t")
normal$case <- "normal case"
#special$case <- "special case"
pooled$case <- "pooled, normal and special cases"
d<-rbind(normal,pooled)
d<-d %.% 
  filter(kanton=="CH",steuerperiode==1993.5)

@@

<<polindexspecialcases9394, results='asis'>>=
polTable(pooled_i,normal_i)
@

1993/94 a pooled data set of normal and special cases has a slightly higher density at both ends compared to data based in normal cases only. Special cases appear to have a slightly lower median income and their distribution is more skewed. Therefore special cases are more polarized than normal cases, i.e. striving away from the median (positive Median Index of 0.02). This tendency is more pronounces in the lower than upper part  of the distribution (Lower Index of 0.029 compared to Upper Index if 0.01). 
As the special cases consist of a broad mix of individuals it remains unclear which factor explains the differences of both distributions. Possible explanations can be immigrants partly concentrating in lower income percentiles, low income artists who belong to the special cases or a more technical selection effect: tax units not liable for taxation throughout the whole year are special cases; those cases might have lower incomes, e.g. if they moved and stopped working. To get a more complete picture we can look how the two distributions relate to each other in 2010 (see FIGURE SPECIAL2010).

% Variante 2010
<<setup_data_specialcases10>>=
library(foreign)
library(ggplot2)
#library(reldist)
normal <- filter(read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t"),kanton=="CH",steuerperiode==2010)
normal <- select(normal, p1:p99, kanton, steuerperiode, -p96,-p97,-p98)
start <- which(names(normal)=="p1")
end <- which(names(normal)=="p99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,99)
normal <- data.frame(t(normal[1,start:end]),percentile)
names(normal)[1]=c("inc")

@@

<<polindexspecialcases10, results='asis'>>=
polTable(pooled_i,normal_i)
@

@@

<<with_without_zeros_Gini,warning=FALSE, fig.cap='',fig.pos='H'>>=
df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
library(reshape)
dflong <- melt(df[,c("G_steink","G_steink0","G_steink_halb","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
names(dflong)[3]<-"Type"
names(dflong)[4]<-"Gini"
levels(dflong$Type)<-c("without zeros","including zeros","imputed values for zeros")
ggplot(dflong[dflong$kanton=="CH" & dflong$steuerperiode>=1995.5,], 
       aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+
  scale_y_continuous(limits=c(0.18,0.6))+
  scale_x_continuous(limits=c(1994,2012),breaks=number_ticks(10)) +
  geom_line(size=0.5)+
  geom_point(aes(shape=Type),size=3)+
  theme_bw()+
  xlab("tax period")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))

@@

<<polindexzeros, results='asis'>>=
polTable(bd.2010.ind,bd.0.2010.ind)
@

@@

<<polindexzeros2, results='asis'>>=
polTable(bd.2010.ind,bd.0.2010.ind)
@

@@

<<setup_data Brülhart_mitNuller_und_HABE_2010,warning=FALSE>>=
library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
library(foreign)
bd.2010 <- filter(read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
start <- which(names(bd.2010)=="p1")
end <- which(names(bd.2010)=="p99_99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
names(bd.2010)[1]=c("inc")
bd.2010 <- rbind(c(0,0),bd.2010)
ival<-bd.2010$inc
perval<-bd.2010$percentile
bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
  len=(perval[i]-perval[i-1])*100
  seq(ival[i-1],ival[i],length.out=len)
}))

@@

#####################################################
library(reshape, quietly=TRUE, warn.conflicts=FALSE)
library(dplyr, quietly=TRUE, warn.conflicts=FALSE)
try(habe0911<-read.table("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
try(habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
try(habe0911<-read.table("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
try(g2009 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2009_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2010 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2010_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2011 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2011_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2009 <- read.csv("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE2009_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2010 <- read.csv("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE2010_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2011 <- read.csv("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE2011_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2009 <- read.csv("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE2009_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2010 <- read.csv("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE2010_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2011 <- read.csv("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE2011_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
names(g2009)[3:4]<-c("SRH","Gewicht")
names(g2010)[3:4]<-c("SRH","Gewicht")
names(g2011)[3:4]<-c("SRH","Gewicht")
jg091011 <- rbind(g2009,g2010,g2011)
#all.equal(match(jg091011$HaushaltID,habe0911$HaushaltID),1:nrow(habe0911)) ## proves it is ok to just attach weights in the default order
habe0911$jahresgewicht <- jg091011$Gewicht

@@

<<BruelhartvsHABE,warning=FALSE, fig.cap='',fig.pos='H'>>=
library(ggplot2)
library(reldist, quietly=TRUE, warn.conflicts = FALSE)

@@

par(mfrow=c(1,2))
# Probability density Function (FTA (Brülhart vs HBS Brutto (korrigiert)))
density1 <- density(bd.2010.ind)
plot(x = (density1$x), y = density1$y, type = "l",
     xlab = "Einkommen", ylab = "density",
     axes = FALSE,
     xlim = c(-30, 800))
     #ylim=c(0,1.151e-02 ))
title(main="FTA vs HBS income distribution",cex=0.6)
axis(side = 1)
axis(side = 2)
fig1legend <- list(x=c(400,400),y=c(0.004,0.004))
legend(fig1legend,lty=1:2,cex=1, bty="n",
       legend=c("FTA","HBS"))
density2 <- density(habe10$psteink,weights=habe10$gew/sum(habe10$gew))
lines(x = (density2$x), y = density2$y, type = "l",lty=2)

@@

par(mfrow=c(1,2))
density1 <- density(bd.2010.ind)
plot(x = (density1$x), y = density1$y, type = "l",
     xlab = "Einkommen", ylab = "density",
     axes = FALSE,
     xlim = c(min(rbind(density1$x,density2$x)), 800),
     ylim=c(0,max(rbind(density1$y,density2$y))))
title(main="FTA vs HBS income distribution",cex=0.6)
axis(side = 1)
axis(side = 2)
fig1legend <- list(x=c(median(rbind(density1$x,density2$x)),median(rbind(density1$x,density2$x))),y=c(max(rbind(density1$y,density2$y)),max(rbind(density1$y,density2$y))))
legend(fig1legend,lty=1:2,cex=1, bty="n",
       legend=c("HBS","FTA"))
density2 <- density(habe10.married$psteink,weights=habe10.married$gew/sum(habe10.married$gew))
lines(x = (density2$x), y = density2$y, type = "l",lty=2)

@@

par(mfrow=c(1,2))
density1 <- density(habe10.unmarried$inc_habe)
plot(x = (density1$x), y = density1$y, type = "l",
     xlab = "Einkommen", ylab = "density",
     axes = FALSE,
     xlim = c(min(rbind(density1$x,density2$x)), 800),
     ylim=c(0,max(rbind(density1$y,density2$y))))
title(main="FTA vs HBS income distribution",cex=0.6)
axis(side = 1)
axis(side = 2)
fig1legend <- list(x=c(median(rbind(density1$x,density2$x)),median(rbind(density1$x,density2$x))),y=c(max(rbind(density1$y,density2$y)),max(rbind(density1$y,density2$y))))
legend(fig1legend,lty=1:2,cex=1, bty="n",
       legend=c("HBS","FTA"))
density2 <- density(bd.2010.ind)
lines(x = (density2$x), y = density2$y, type = "l",lty=2)

@@

<<benplot, fig.cap='',fig.pos='H'>>=
 ### libs
library(reshape, quietly=TRUE, warn.conflicts = FALSE)
library(foreign)
library(dplyr)
library(reldist)
library(ggplot2)

>>>>

<<libraries, include=FALSE>>=
library(foreign,quietly=TRUE, warn.conflicts=FALSE)
library(ggplot2)
library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
library(reshape)
# before loading reldist we load nlme and mgcv. else reldist prints library loading messages
library(nlme,quietly=TRUE, warn.conflicts=FALSE)
library(mgcv,quietly=TRUE, warn.conflicts=FALSE)
library(reldist,quietly=TRUE,warn.conflicts=FALSE) #for reldist plots and inequality measures
library(Hmisc,quietly=TRUE,warn.conflicts=FALSE) #for polTable function / Latex tables
@

@@

<<setup_gini_data>>=
df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
df$zeros <- df$null_norm/df$cpop # i only use normal cases here
@

@@

<<setup_data_Brühlhart with and without taxed>>=
# Ohne Nuller
bd <- read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd$Jahr <- as.numeric(substr(bd$Veranlagungsperiode,1,4))
bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] <- bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] - 2
bd<-bd %.% 
  filter(Einheit=="Total")
bd<-subset(bd,Jahr<1995 | Jahr >2002)
bd.long<-reshape(bd,
varying = c("gini_reink","gini_reinka"),
v.names = "Gini",
timevar = "inc_def",
times = c("Net income (only taxed)","Equivalised net income (only taxed)"),
direction ="long")
# mit Nuller
bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
bd.0<-bd.0 %.% 
  filter(Einheit=="Total")
bd.0<-subset(bd.0,Jahr<1993 | Jahr >2002)
bd.0.long<-reshape(bd.0,
varying = c("gini_reink","gini_reinka"),
v.names = "Gini",
timevar = "inc_def",
times = c("Net income","Equivalised net income"),
direction ="long")
bdlong<-rbind(bd.0.long,bd.long)
names(bdlong)[61]<-"Income"
@

<<equivalencescale, fig.cap='',fig.pos='H'>>=
number_ticks <- function(n) {function(limits) pretty(limits, n)}
ggplot(bdlong, aes(x=Jahr,y=Gini,shape=Income))+
  scale_y_continuous(limits=c(0.3,0.55))+
  scale_x_continuous(limits=c(1970,2012),breaks=number_ticks(10)) +
  geom_point(aes(shape=Income),size=3)+
  geom_line(data=bdlong,aes(linetype=Income))+
  theme_bw()+
  xlab("tax period")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))
@

@@

<<setup_data_Brülhart_2003>>=
bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
bd.2003<-bd.0 %.% 
  filter(Jahr==2003,Einheit=="Total")
start <- which(names(bd.2003)=="p1")
end <- which(names(bd.2003)=="p99_99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
bd.2003<- data.frame(t(bd.2003[1,start:end]),percentile)
names(bd.2003)[1]=c("inc")
cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
bd.2003 <- rbind(c(0,0),bd.2003)
ival<-bd.2003$inc
perval<-bd.2003$percentile
bd.2003.ind<-unlist(sapply(2:length(ival),function(i){
  len=(perval[i]-perval[i-1])*100
  seq(ival[i-1],ival[i],length.out=len)
}))
bd.2003.ind<-bd.2003.ind/1000

@@

<<setup_data_Brülhart_2010>>=
bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
bd.2010<-bd.0 %.% 
  filter(Jahr==2010,Einheit=="Total")
start <- which(names(bd.2010)=="p1")
end <- which(names(bd.2010)=="p99_99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
names(bd.2010)[1]=c("inc")
bd.2010 <- rbind(c(0,0),bd.2010)
ival<-bd.2010$inc
perval<-bd.2010$percentile
bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
  len=(perval[i]-perval[i-1])*100
  seq(ival[i-1],ival[i],length.out=len)
}))
bd.2010.ind<-bd.2010.ind/1000
@

<<reldist20032010, echo=FALSE, fig.cap='',fig.pos='H'>>=

@@

<<poltable>>=
# Setup functin to produce latex tables with polarization indices and difference in ginis
polTable <- function(y, yo, ywgt=1, yowgt=1,ndig=2,title="Inequality Indices",caption='',tablabel) {
stopifnot(is.numeric(y), is.numeric(yo),exists("tablabel")) # all TRUE
#y vector with new distribution
#yo vector with reference distribution
#ywgt yowgt: weights to use (optional)
#ndig: digits to display, default 2
#title: table title (optional)

@@

latex(format(c("Median Index"=med, "Lower Index"=lower, "Upper Index"=upper, "$\\Delta$ Gini"=dgini),digits=ndig),file="",title=title,where='H',label=tablabel,caption=caption)
}

<<polindex20032010, results='asis'>>=
polTable(bd.2010.ind,bd.2003.ind, tablabel="pol20032010")
@

@@

% Variante 1993/1994
<<setup_data_ESTVspecialcases9394>>=
#library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
#library(ggplot2, quietly=TRUE, warn.conflicts = FALSE)

normal <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
normal <- select(normal, p1:p99, kanton, steuerperiode, -p96,-p97,-p98)
#normal <- normal %.% select(-G_reink, -G_taxed, -ppop0, -G_steink0)
pooled <- read.csv("../../data/ginis_und_perzentile_beides.csv",sep="\t")
pooled <- select(pooled, kanton, steuerperiode, p1:p99)
#special <- read.csv("data/ginis_und_perzentile_sonder.csv",segep="\t")
normal$case <- "normal case"
#special$case <- "special case"
pooled$case <- "pooled, normal and special cases"
d<-rbind(normal,pooled)
d<-d %.% 
  filter(kanton=="CH",steuerperiode==1993.5)

@@

<<polindexspecialcases9394, results='asis'>>=
polTable(pooled_i,normal_i,tablabel="pol9394spec")
@

1993/94 a pooled data set of normal and special cases has a slightly higher density at both ends compared to data based in normal cases only (see figure \ref{fig:specialcases9394}). Special cases appear to have a slightly lower median income and their distribution is more skewed. Therefore special cases are more polarized than normal cases (see table \ref{pol9394spec}), i.e. striving away from the median (positive Median Index of 0.02). This tendency is more pronounced in the lower than upper part  of the distribution (Lower Index of 0.029 compared to Upper Index if 0.01). 
As the special cases consist of a broad mix of individuals it remains unclear which factor explains the differences of both distributions. Possible explanations can be immigrants partly concentrating in lower income percentiles, low income artists who belong to the special cases or a more technical selection effect: tax units not liable for taxation throughout the whole year are special cases; those cases might have lower incomes, e.g. if they moved and stopped working. To get a more complete picture we can look how the two distributions relate to each other in 2010 (see figure \ref{fig:specialcases10}).

% Variante 2010
<<setup_data_specialcases10>>=
#library(foreign)
#library(ggplot2)
#library(reldist)
normal <- filter(read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t"),kanton=="CH",steuerperiode==2010)
normal <- select(normal, p1:p99, kanton, steuerperiode, -p96,-p97,-p98)
start <- which(names(normal)=="p1")
end <- which(names(normal)=="p99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,99)
normal <- data.frame(t(normal[1,start:end]),percentile)
names(normal)[1]=c("inc")

@@

<<polindexspecialcases10, results='asis'>>=
polTable(pooled_i,normal_i,tablabel="pol2010spec")
@

@@

<<with_without_zeros_Gini,warning=FALSE, fig.cap='',fig.pos='H'>>=
df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
#library(reshape)
dflong <- melt(df[,c("G_steink","G_steink0","G_steink_halb","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
names(dflong)[3]<-"Type"
names(dflong)[4]<-"Gini"
levels(dflong$Type)<-c("without zeros","including zeros","imputed values for zeros")
ggplot(dflong[dflong$kanton=="CH" & dflong$steuerperiode>=1995.5,], 
       aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+
  scale_y_continuous(limits=c(0.18,0.6))+
  scale_x_continuous(limits=c(1994,2012),breaks=number_ticks(10)) +
  geom_line(size=0.5)+
  geom_point(aes(shape=Type),size=3)+
  theme_bw()+
  xlab("tax period")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))

@@

<<polindexzeros, results='asis'>>=
polTable(bd.2010.ind,bd.0.2010.ind,tablabel="polzeros")
@

@@

<<polindexzeros2, results='asis'>>=
polTable(bd.2010.ind,bd.0.2010.ind,tablabel="polzeros2")
@

@@

<<setup_data Brülhart_mitNuller_und_HABE_2010,warning=FALSE>>=
#library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
#library(foreign)
bd.2010 <- filter(read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
start <- which(names(bd.2010)=="p1")
end <- which(names(bd.2010)=="p99_99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
names(bd.2010)[1]=c("inc")
bd.2010 <- rbind(c(0,0),bd.2010)
ival<-bd.2010$inc
perval<-bd.2010$percentile
bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
  len=(perval[i]-perval[i-1])*100
  seq(ival[i-1],ival[i],length.out=len)
}))

@@

#####################################################
#library(reshape, quietly=TRUE, warn.conflicts=FALSE)
#library(dplyr, quietly=TRUE, warn.conflicts=FALSE)
try(habe0911<-read.table("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
try(habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
try(habe0911<-read.table("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
try(g2009 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2009_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2010 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2010_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2011 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2011_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2009 <- read.csv("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE2009_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2010 <- read.csv("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE2010_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2011 <- read.csv("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE2011_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2009 <- read.csv("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE2009_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2010 <- read.csv("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE2010_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2011 <- read.csv("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE2011_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
names(g2009)[3:4]<-c("SRH","Gewicht")
names(g2010)[3:4]<-c("SRH","Gewicht")
names(g2011)[3:4]<-c("SRH","Gewicht")
jg091011 <- rbind(g2009,g2010,g2011)
#all.equal(match(jg091011$HaushaltID,habe0911$HaushaltID),1:nrow(habe0911)) ## proves it is ok to just attach weights in the default order
habe0911$jahresgewicht <- jg091011$Gewicht

@@

<<BruelhartvsHABE,warning=FALSE, fig.cap='',fig.pos='H'>>=
#library(ggplot2)
#library(reldist, quietly=TRUE, warn.conflicts = FALSE)

@@

par(mfrow=c(1,2))
# Probability density Function (FTA (Brülhart vs HBS Brutto (korrigiert)))
density1 <- density(bd.2010.ind)
plot(x = (density1$x), y = density1$y, type = "l",
     xlab = "income", ylab = "density",
     axes = FALSE,
     xlim = c(-30, 800))
     #ylim=c(0,1.151e-02 ))
title(main="FTA vs HBS income distribution",cex=0.6)
axis(side = 1)
axis(side = 2)
fig1legend <- list(x=c(400,400),y=c(0.004,0.004))
legend(fig1legend,lty=1:2,cex=1, bty="n",
       legend=c("FTA","HBS"))
density2 <- density(habe10$psteink,weights=habe10$gew/sum(habe10$gew))
lines(x = (density2$x), y = density2$y, type = "l",lty=2)

@@

par(mfrow=c(1,2))
density1 <- density(bd.2010.ind)
plot(x = (density1$x), y = density1$y, type = "l",
     xlab = "income", ylab = "density",
     axes = FALSE,
     xlim = c(min(rbind(density1$x,density2$x)), 800),
     ylim=c(0,max(rbind(density1$y,density2$y))))
title(main="FTA vs HBS income distribution",cex=0.6)
axis(side = 1)
axis(side = 2)
fig1legend <- list(x=c(median(rbind(density1$x,density2$x)),median(rbind(density1$x,density2$x))),y=c(max(rbind(density1$y,density2$y)),max(rbind(density1$y,density2$y))))
legend(fig1legend,lty=1:2,cex=1, bty="n",
       legend=c("HBS","FTA"))
density2 <- density(habe10.married$psteink,weights=habe10.married$gew/sum(habe10.married$gew))
lines(x = (density2$x), y = density2$y, type = "l",lty=2)

@@

par(mfrow=c(1,2))
density1 <- density(habe10.unmarried$inc_habe)
plot(x = (density1$x), y = density1$y, type = "l",
     xlab = "income", ylab = "density",
     axes = FALSE,
     xlim = c(min(rbind(density1$x,density2$x)), 800),
     ylim=c(0,max(rbind(density1$y,density2$y))))
title(main="FTA vs HBS income distribution",cex=0.6)
axis(side = 1)
axis(side = 2)
fig1legend <- list(x=c(median(rbind(density1$x,density2$x)),median(rbind(density1$x,density2$x))),y=c(max(rbind(density1$y,density2$y)),max(rbind(density1$y,density2$y))))
legend(fig1legend,lty=1:2,cex=1, bty="n",
       legend=c("HBS","FTA"))
density2 <- density(bd.2010.ind)
lines(x = (density2$x), y = density2$y, type = "l",lty=2)

@@

<<benplot, fig.cap='',fig.pos='H'>>=
 ### libs
#library(reshape, quietly=TRUE, warn.conflicts = FALSE)
